Internal models of type theory and internal higher categories for the Univalent Foundations

### 2014-03-27 article-models-ttmodels.tex
figuring out what families of extended models are

We consider the following diagram helpful:
\begin{equation*}
\begin{tikzcd}
\jcomp{}{\bar{\typefont{pr}}}{{}{\tfext{A}}{\mfam{P}}}
  \rar{}
  \ar[fib]{d}
  &
  \jcomp{}{\tfext{A}}{\mfam{P}}
    \rar{}
    \ar[fib]{d}{}
  &
  \mfam{P}
    \ar[fib]{d}{}
  \\
\jcomp{}{\bar{\typefont{pr}}}{{}{\tfext{A}}{P}}
  \rar{}
  \ar[fib]{d}{}
  &
  \jcomp{}{\tfext{A}}{P}
  \rar{}
  \ar[fib]{d}{}
  &
  P
  \ar[fib]{d}{}\\
\ctxext{{A}{P}}{\mfam{A}}
  \ar{r}[swap]{\bar{\typefont{pr}}}
  &
  \ctxext{A}{\mfam{A}}
  \ar{r}[swap]{\tfext{A}}
  &
  A
\end{tikzcd}
\end{equation*}
Here $\bar{\typefont{pr}}\jdeq\tmext{}{}{\cprojfstf{A}{P}}{\idtm{\mfam{A}}}$,
the unmarked horizontal arrows indicate identity terms and the vertical arrows
merely indicate that what's above is a family depending on what's below.

>>>>

We consider the following dependency diagram helpful:
\begin{equation*}
\begin{tikzcd}
\jcomp{}{\bar{\typefont{pr}}}{{}{\tfext{A}}{\mfam{P}}}
  \ar[fib]{d}
  &
  \jcomp{}{\tfext{A}}{\mfam{P}}
    \ar[fib]{d}{}
  &
  \mfam{P}
    \ar[fib]{d}{}
  \\
\jcomp{}{\bar{\typefont{pr}}}{{}{\tfext{A}}{P}}
  \ar[fib]{d}{}
  &
  \jcomp{}{\tfext{A}}{P}
  \ar[fib]{d}{}
  &
  P
  \ar[fib]{d}{}\\
\ctxext{{A}{P}}{\ctxwk{P}{\mfam{A}}}
  \ar{r}[swap]{\bar{\typefont{pr}}}
  &
  \ctxext{A}{\mfam{A}}
  \ar{r}[swap]{\tfext{A}}
  &
  A
\end{tikzcd}
\end{equation*}
Here the morphism 
$\bar{\typefont{pr}}$ should be explained (skip ahead to 
\autoref{barproj} for the definition). Note that we have the judgmental
equalities
\begin{align*}
\ctxwk{\ctxext{{A}{P}}{\ctxwk{P}{\mfam{A}}}}{\ctxext{A}{\mfam{A}}}
& \jdeq 
  \ctxext
    {\ctxwk{\ctxext{{A}{P}}{\ctxwk{P}{\mfam{A}}}}{A}}
    {\ctxwk{\ctxext{{A}{P}}{\ctxwk{P}{\mfam{A}}}}{\mfam{A}}}
  \\
& \jdeq
  \ctxext
    {\ctxwk{{P}{\mfam{A}}}{{\ctxext{A}{P}}{A}}}
    {\ctxwk{\ctxext{{A}{P}}{\ctxwk{P}{\mfam{A}}}}{\mfam{A}}}
\end{align*}
Note that we have the term $\ctxwk{{P}{\mfam{A}}}{\cprojfstf{A}{P}}$ of the
family $\ctxwk{{P}{\mfam{A}}}{{\ctxext{A}{P}}{A}}$. Therefore, we need to
find a term of type $\subst{\ctxwk{{P}{\mfam{A}}}{\cprojfstf{A}{P}}}
{\ctxwk{\ctxext{{A}{P}}{\ctxwk{P}{\mfam{A}}}}{\mfam{A}}}$. Note that we have
the judgmental equalities:
\begin{align*}
\subst
  {\ctxwk{{P}{\mfam{A}}}{\cprojfstf{A}{P}}}
  {\ctxwk{\ctxext{{A}{P}}{\ctxwk{P}{\mfam{A}}}}{\mfam{A}}}
& \jdeq
  \subst
    {\ctxwk{{P}{\mfam{A}}}{\cprojfstf{A}{P}}}
    {\ctxwk{{P}{\mfam{A}}}{{\ctxext{A}{P}}{\mfam{A}}}}
  \\
& \jdeq
  \ctxwk
    { {P}{\mfam{A}}
      }
    { \subst
        {\cprojfstf{A}{P}}
        {\ctxwk{\ctxext{A}{P}}{\mfam{A}}}
      }
  \\
& \jdeq
  \ctxwk
    { {P}{\mfam{A}}
      }
    { {P}{\mfam{A}}
      }
  \\
& \jdeq
  \ctxwk{P}{{\mfam{A}}{\mfam{A}}}
\end{align*}
We find the term $\ctxwk{P}{\idtm{\mfam{A}}}$ here. Thus we can now define
$\bar{\typefont{pr}}$ by:
\begin{equation}\label{barproj}
\jhomdefn
  {\Gamma}
  {{{A}{P}}{\mfam{A}}}
  {{A}{\mfam{A}}}
  {\bar{\typefont{pr}}}
  {\tmext{}{}{\ctxwk{{P}{\mfam{A}}}{\cprojfstf{A}{P}}}{\ctxwk{P}{\idtm{\mfam{A}}}}}
\end{equation}

\begin{lem}
We have the judgmental equality
\begin{equation*}
\jfameq
  {{{{\Gamma}{A}}{P}}{\ctxwk{P}{\mfam{A}}}}
  {\jcomp{}{\bar{\typefont{pr}}}{Q}}
  {\ctxwk{P}{Q}}
\end{equation*}
for any family $Q$ of contexts over $\ctxext{{\Gamma}{A}}{\mfam{A}}$ 
\end{lem}
